---
title: "Coursera Data Science Capstone Project Week 2 Assignment"
author: "Johnny Kao"
date: "April 20, 2019"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(stringi)
library(ggplot2)
library(tm)
# library(RWeka)
```

## Basic Information

The goal of this project is just to display that you've gotten used to working with the data and that you are on track to create your prediction algorithm. Please submit a report on R Pubs <http://rpubs.com/> that explains your exploratory analysis and your goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. You should make use of tables and plots to illustrate important summaries of the data set.

The motivation for this project is to:

1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

## 1. Download and load the files

### Download data files, unzip to working directory -
<https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip>

### Load files
```{r Loading data into R, warning=FALSE, cache=TRUE}
setwd("~/Downloads/final/en_US")

# News
con <- file("en_US.news.txt", open="r")
news <- readLines(con); close(con)

# Blog
con <- file("en_US.blogs.txt", open="r")
blogs <- readLines(con); close(con) 

# Twitter
con <- file("en_US.twitter.txt", open="r")
twitter <- readLines(con); close(con)
```

## 2. Statistics of datasets

We already loaded 3 files, here is the basic information:

### Blogs

```{r File infor - Blogs, cache=TRUE}
# Analysing lines and characters of Blogs
stri_stats_general(blogs)
```

```{r Summary Corpus - Blogs, cache=TRUE, warning=FALSE}
blog_words <- stri_count_words(blogs); summary(blog_words); qplot(blog_words, bins = 50)
```

### Twitter

```{r File infor - Twitter, cache=TRUE}
# Analysing lines and characters of Blogs
stri_stats_general(twitter)
```

```{r Summary Corpus - Twitter, cache=TRUE, warning=FALSE}
twit_words <- stri_count_words(twitter); summary( twit_words); qplot(twit_words, bins = 50)
```

### News

```{r File infor - News, cache=TRUE}
# Analysing lines and characters of Blogs
stri_stats_general(news)
```

```{r Summary Corpus - News, cache=TRUE, warning=FALSE}
news_words <- stri_count_words(news); summary(news_words); qplot(news_words, bins = 50)
```

## 3. Futher analysis of the Corpus

### Sampling the corpus for further analysis
```{r Sampling - Blogs, cache=TRUE, warning=FALSE}
set.seed(1000)
# Sampling of blogs
sampleBlogs <- blogs[sample(1:length(blogs), 1000)]; head(sampleBlogs, 3)

# Sampling of twitter
sampleTwitter <- twitter[sample(1:length(twitter), 1000)]; head(sampleTwitter,3)

# Sampling of news
sampleNews <- news[sample(1:length(news), 1000)]; head(sampleNews, 3)

# Conslidating the sample files
sampleData <- c(sampleTwitter,sampleNews,sampleBlogs)
# Removing the original datasets from memory
rm(blogs); rm(twitter); rm(news)
```

### Cleaning the Corpus
```{r Cleaning the Corpus, cache=TRUE, warning=FALSE}

myCorpus <- Corpus(VectorSource(sampleData))
toSpace  <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
myCorpus <- tm_map(myCorpus, toSpace,"\"|/|@|\\|")
myCorpus <- tm_map(myCorpus, content_transformer(tolower))
myCorpus <- tm_map(myCorpus, removeNumbers)
myCorpus <- tm_map(myCorpus, stripWhitespace)

# Remove stop words
myCorpus <- tm_map(myCorpus , removeWords, stopwords('english'))

```

### Creating ngrams
```{r Creating ngrams, cache=TRUE, warning=FALSE}

myCorpusDF <-data.frame(text = unlist(sapply(myCorpus, `[`, "content")), 
                        stringsAsFactors = FALSE)

findNGrams <- function(corp, grams, top) {
        ngram <- NGramTokenizer(corp, Weka_control(min = grams, max = grams,
                                                   delimiters = " \\r\\n\\t.,;:\"()?!"))
        ngram <- data.frame(table(ngram))
        ngram <- ngram[order(ngram$Freq, decreasing = TRUE),][1:top,]
        colnames(ngram) <- c("Words","Count")
        ngram
        }



```



## What's Next?

The next steps in the project are:

1. Continuing cleaning the corpus to to increase the accuracy of the model
2. Refining the sampling process for getting a good ngram representation without using the entire corpus
3. Building the final prediction model and testing it

For predicting the current word, the application will use mono-gram prediction data. For predicting the next word, the application will use bi-gram, tri-gram and quadri-gram prediction data.
